---
title: "Analyzing Data from Pew's 2017 Global Attitudes Survey"
author: "Rebecca Smith"
date: "3/24/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# INTRODUCTION
For this project, I will be doing another "guided" tutorial provided by medium.com (reference below). I've chosen to do this because I want to learn how to use the CRAN package "pewmethods". 

I will be analyzing data from the Pew Research Center's spring 2017 Global Attitudes Survey, a poll that asks respondents in multiple countries about issues ranging from politics to economic conditions.

# DOWNLOAD AND INSTALL LIBRARIES
First, I will install the devtools package which will allow me to pull the pewmethods package from Pew's github repo. For the subsequent analyses, I will also need tidyverse, haven, rworldmap, and ggrepel packages. 

```{r install}
library(devtools)
install_github("pewresearch/pewmethods", build_vignettes = TRUE, force = TRUE)
library(pewmethods)
library(tidyverse)
library(rworldmap)
library(ggrepel)
```

# DOWNLOAD DATASET
Next, I will import the Global Attitudes Spring 2017 Dataset and save it as "gap_rd". This dataset contains more than 800 variables and 40,000 observations.

```{r data}
library(haven)
gap_rd <- haven::read_sav("Pew-Research-Center-Spring-2017-Global-Attitudes-Survey_September-2018/Pew Research Global Attitudes Spring 2017 Dataset WEB FINAL.sav")
gap_rd
```

# SELECT VARIABLES OF INTEREST

For this project, we will examine two economically focused questions from the survey that were published in the report, "Global Publics More Upbeat About the Economy" (reference below). Because we will just be looking at two questions, I will first narrow down the dataset to the variables of interest: country (country variable), econ_sit (current economic situation in country), children_betteroff2 (will children be better off/worse off than parents), and weight (survey weights). 

```{r select}
gap2017 <- gap_rd %>%
  select(Country, econ_sit, children_betteroff2, weight) %>%
  haven::as_factor()
```

# TWO QUESTIONS

Now, we will explore two questions: 
1. How did people view their country's economic situation in 2017?
2. Did people think that children in their country would be better or worse off than their parents?

First, let's identify the countries included in the survey, find the exact wording of the questions based on the SPSS labels, and get country-level estimates for these two questions of interest.

```{r countries included}
unique(as.character(gap2017$Country))
```
38 countries were included in this dataset.
```{r spss label}
get_spss_label(gap2017, c("econ_sit", "children_betteroff2"))
```
Alright, now we know that the econ_sit variable corresponds to Q5 which asks, "Now thinking about our economic situation, how would you describe the current economic situation in (survey country) - is it very good, somewhat good, somewhat bad or very bad?". The children_betteroff2 variables corresponds to Q8, which asks, "When children today in (survey country) grow up, do you think they will be better off or worse off financially than their parents?".

#CONVERT DATA

The pewmethods package allows us to convert the data from individual - to country-level data. The code below instructs R to look at the econ_sit variable, apply survey weights (weight), group the data by the country2 variable, and round it to the nearest whole number. The result will be in a wide format.

```{r convert_wide}
econ_data_wide <- get_totals("econ_sit", gap2017, wt = "weight", by = "Country", na.rm = T, digits = 0)
head(econ_data_wide)
```

The tutorial suggests that having the data in a long format, rather than wide, may help with analytic and graphing purposes. Below is the code I will use to attempt that:

```{r convert_long}
econ_data_long <- econ_data_wide %>%
  gather(country, economy, -one_of("econ_sit", "weight_name"), factor_key = T)
head(econ_data_long)
```

The tutorial explains that the team at Pew Research Center that examines this survey data typically collapses the responses into "good" and "bad" categories by adding rounded totals for very/somewhat good and very/somewhat bad. The code to replicate this is below. The spread function will be used to shape the dataset from long back to wide.

```{r collapse}
econ_sit_collapse <- econ_data_long %>%
  mutate(econ_eval = case_when(
    econ_sit == "Very good" ~ "Good",
    econ_sit == "Somewhat good" ~ "Good",
    econ_sit == "Somewhat bad" ~ "Bad",
    econ_sit == "Very bad" ~ "Bad",
    TRUE ~ "DK/Refused"
  )) %>%
  group_by(country, econ_eval) %>%
  summarise(share = sum(economy))

final_econ <- econ_sit_collapse %>% spread(econ_eval, share)
head(final_econ[c("country", "Good", "Bad", "DK/Refused")])
```

# RECREATE MAP
The 2017 report includes a map (reference below) of the data that shows the share of respondents in each country who said the economic situation was good. Here is where the rworldmap package comes into play. Following the tutorial, I am going to try to recreate this map using said package.

First, I need to recode the economy data into categories to mimic the color schme in the report map (though, this time, I will use a green/blue scale), turn it into a factor, and then re-elevel everything to keep it in the desired order (rather than in the default alphabetical order). 

```{r recode econ_sit into cat vars}
final_econ <- final_econ %>%
  mutate(econ_good_cat = case_when(
    Good < 20 ~ "<20%",
    Good %in% 20:34 ~ "20-34",
    Good %in% 35:49 ~ "35-49",
    Good %in% 50:64 ~ "50-64",
    Good %in% 65:80 ~ "60-80",
    Good > 80 ~ ">80%") %>%
  fct_relevel("<20%", "20-34", "35-49", "50-64", "65-80", ">80%")
  ) %>%
  ungroup

sPDF <- joinCountryData2Map(final_econ, joinCode = "NAME", nameJoinColumn = "country")

mapCountryData(sPDF,
               nameColumnToPlot = "econ_good_cat",
               catMethod = "categorical",
               colourPalette = c("#1F497D", "#8DB5E2", "#C5D9F1", "#C4CA8A", "#949D48", "#707636"),
               mapTitle = "% who say the current economic situation in our country is good, 2017",
               missingCountryCol = "#F2F2F2")
```

YAY! It worked! The resulting map mimics the map included in the report. 

# LOOKING AT CHILDREN BETTER/WORSE OFF IN EUROPEAN COUNTRIES
The survey includes a question that asks whether respondents think their children will be better or worse off financially than their parents. Let's look at the positive and negative sides of the question for some European countries surveyed.

First, I have to filter the dataset.

```{r filter euro}
gap2017_EU <- gap2017 %>%
  filter(Country %in% c("United Kingdom", "France", "Germany", "Netherlands", "Spain", "Sweden", "Poland", "Greece", "Hungary", "Italy"))

children_data_wide <- get_totals("children_betteroff2",
                                 gap2017_EU, wt = "weight", by = "Country", na.rm = T, digits = 0)

children_data_long <- children_data_wide %>% gather(country, value, -one_of("children_betteroff2", "weight_name"), factor_key = T)

head(children_data_long)
sample_n(children_data_long, 10)
```

# CREATING BAR CHARTS

One of the reasons why I picked this tutorial is because I want to learn how to create accurate, succinct, and visually-appealing graphs to represent Pew data. This tutorial explains that the global attitudes team at Pew often displays the data in horizontal bar charts. To do this, we first need to manipulate the data in order to make a plot similar to those published in Pew reports. The bar chart will require both negative and positive values, so I will first create a new variable called children_bars that makes any "worse off" category negative while keeping the "better off" values positive. I will then mad the bar color to our graphic using "fill = children_betteroff2". I can then use the new children_bars variable to sort the data based on the "better off" values. 

To create the dataset I need for the graphic, I will filter the children_data_long dataframe to only the "better off" or "worse off" responses, pipe it to ggplot2 and customize the design of the graphic. The fct_reorder function will allow me to sort the data by the percent who say "better off" at the country level in descending order.

```{r horizontal bargraph}
#specify top margin
margin(t = 10)
p <- children_data_long %>%
  filter(!(children_betteroff2 %in% c("Don’t know (DO NOT READ)",
                                      "Same (DO NOT READ)",
                                      "Refused (DO NOT READ)")
  )) %>%
  mutate(
    children_bars = ifelse(children_betteroff2 == "Worse off", value * -1, value),
    children_betteroff2 = fct_relevel(children_betteroff2, 
                                      "Worse off", "Better off"),
    better_only = ifelse(children_betteroff2 == "Better off", value, 0)
  ) %>% 
  ggplot(
    aes(x = reorder(country, better_only),
        y = children_bars,
        fill = children_betteroff2,
        group = children_betteroff2
    )
  ) +
  geom_bar(stat = "identity") +
  geom_label(
    aes(
      label = abs(as.numeric(children_bars))
      
    ),
    fill = "white"
  ) + 
  scale_fill_manual(values = c("Better off" = "#949D48", "Worse off" = "#006699")) +
  coord_flip() +
  theme(
    axis.line = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "top", 
    legend.title = element_blank(),
    axis.title = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = 11, face = "bold")
  )
plot(p)
```
From this graphic, we can see that respondents from all countries included in this visualization believe that their children will be worse off financially. Well, that's depressing. AND that wraps up this portfolio piece!

#REFERENCES
Devlin, K. (2020, April 13). Analyzing international survey data with the pewmethods R package. Medium. https://medium.com/pew-research-center-decoded/analyzing-international-survey-data-with-the-pewmethods-r-package-3b0b21cba607

NW, 1615 L. St, Suite 800Washington, & Inquiries, D. 20036USA202-419-4300 | M.-857-8562 | F.-419-4372 | M. (2017, June 5). Global Publics More Upbeat About the Economy. Pew Research Center’s Global Attitudes Project. https://www.pewresearch.org/global/2017/06/05/global-publics-more-upbeat-about-the-economy/

